---
title: "Intro to NTLlakeloads"
author: 'Hilary Dugan'
date: '2020-04-27'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(NTLlakeloads)
library(tidyverse)
```

```{r, include=FALSE}
devtools::load_all(".")
```

## Get LTER data
```{r getLTERdata, eval = T, warning = F, message = F}
# Load NTL datasets
LTERtemp = loadLTERtemp() # Download NTL LTER data from EDI
LTERnutrients = loadLTERnutrients() # Download NTL LTER data from EDI
LTERions = loadLTERions() # Download NTL LTER data from EDI
LTERsecchi = loadLTERsecchi() # Download NTL LTER data from EDI
```

### Additionally, these datasets can be viewed at
### https://github.com/hdugan/NTLviewer

## Variables available for plotting
```{r, eval = FALSE}
# Available variables
availableVars()

# Available variables that are not depth-discrete. Used with weeklyInterpolate.1D.
availableVars.1D()
```

## Interpolate weekly total phosphorus data for Lake Mendota
```{r, eval = T, warning = F, message = F, results = 'hide'}
# printFigs = TRUE to output series of interpolated profiles (but slower)
# See help file for parameter descripts
df.ME = weeklyInterpolate(lakeAbr = 'ME', var = 'totpuf_sloh', dataset = LTERnutrients, maxdepth = 24, 
                          constrainMethod = 'zero', setThreshold = 0.1, printFigs = F)
```

## Plotting entire timeseries 
```{r, eval = T, warning = F, message = F, fig.width = 7}
plotTimeseries(df.interpolated = df.ME$weeklyInterpolated, var = 'totpuf_sloh')
```

## Plot specific year with observed data
```{r, eval = T, warning = F, message = F, fig.width = 7 }
# With observations
plotTimeseries.year(df.interpolated = df.ME$weeklyInterpolated, observations = df.ME$observations,  var = 'totpuf_sloh', chooseYear = 2008)

# Without observations, but adding legend title 
plotTimeseries.year(df.interpolated = df.ME$weeklyInterpolated, var = 'totpuf_sloh', chooseYear = 2008, legend.title = 'TP (µg/L)')
```

## Calculate mass at annual or weekly timescales
```{r, eval = T, results = 'hide', warning = F, message = F}
df.mass.annual = calcMass(df.ME$weeklyInterpolated,lakeAbr = 'ME', time.res = 'annual', conversion = 1e6)
```

## Example of plotting annual mass 
```{r, eval = T, warning = F, message = F, fig.width = 7, fig.height= 4}
library(ggplot2)

ggplot(df.mass.annual, aes(x = year, y = mass)) +
  geom_path() +
  geom_point() +
  ylab('TP (kg)') +
  labs(title = 'Lake Mendota annual TP mass', caption = 'Calculated from NTLlakeloads') +  
  theme_bw(base_size = 10) +
  theme(axis.title.x = element_blank())
```

## Decompose weekly mass timeseries to analyse trends and seasonality
```{r, eval = T, warning = F, message = F}
df.mass = calcMass(df.ME$weeklyInterpolated,lakeAbr = 'ME', time.res = 'weekly', conversion = 1e6)

decomposeTS(df.mass, lakeAbr = 'ME', var = 'totpuf_sloh')
```
